---
title: "Main"
output: html_document
---

## Step 0: Install Libraries
```{r setup, eval=FALSE, warning = FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(Matching)
```

## Step 1:Data

```{r}
smoking_data <- read.csv('nhefshw2dat.csv')
```

## Step 2: Propensity Scoring 

### Step 2.1: Estimating Propensity Scoring with Logistic Regression
```{r}
propensity_model <- glm(qsmk ~ as.factor(sex)+as.factor(race) + age + as.factor(education.code) + smokeintensity + smokeyrs + as.factor(exercise)+ as.factor(active) + wt71, family = binomial() ,data = smoking_data)

summary(propensity_model)
```

Now that we have the logistic regression model created. We are going to estimate the propensity score for each subject in the data. 
```{r}
# Make predictions with the logistic model:
predictions <- predict.glm(propensity_model, smoking_data, type = 'response')
# Create new column in the smoking data:
smoking_data$propensity <- predictions 
```

### Step 2.2 Matching Based On Propensity Scores

Now that we have the estimated propensity scores for each subject we can go ahead and perform matching.
```{r}
library(Matching)
#Perform Matching
x <- smoking_data$propensity
t <- smoking_data$qsmk
y <- smoking_data$wt82_71
rr <- Match(Y = y, Tr = t, X = x, M = 1)
summary(rr)
#Check covariate balance:
matching <- MatchBalance(qsmk ~ as.factor(sex)+as.factor(race) + age + as.factor(education.code) + smokeintensity + smokeyrs + as.factor(exercise)+ as.factor(active) + wt71 ,match.out = rr,data = smoking_data, nboots = 500)
```

As we can see the after matching the covariates are a lot more balanced than they were before. Now we need to find the average absolute mean standardized difference.

There doesn't seem to be a good method to extract the standardized mean difference from the matching function. Therefore, we are going to have to do it manually. 

```{r}
#Check the total aboslute std mean difference before matching:
before_aasmd <- (abs(-16.022) + abs(-19.905) + abs(27.714) + abs(-11.633) + abs(-4.7408) + abs(-2.7616)  + abs(15.215) + abs(-20.874) +  abs(15.26) + abs(3.9669) + abs(5.6429) + abs(2.6753) + abs(7.0522) + abs(13.13))/14

# Check the total absolute std mean difference after matching:
prop_log_total_aasmd <- (abs(.15703) + abs(2.0202) + abs(-3.4504) + abs(-4.4403) + abs(1.7703) + abs(-2.7652) + abs(.56014) + abs(-1.3479) + abs(-3.176) + abs(-4.8493) + abs(7.7689) + abs(-.38737) + abs(4.4974) + abs(-1.3303))/14
```
Now let's compare the average absolute std mean difference in covariates before matching to after matching:
```{r}
#Before Matching:
print(before_aasmd)
#After Matching:
print(prop_log_total_aasmd)
```
Let's visualize the results in a barchart:
```{r}
#Create dataframe to visualize:
aasmd_prop <- data.frame("Type" = c("Before Matching", "After Matching"), "Value" = c(before_aasmd, prop_log_total_aasmd))
library(ggplot2)
aasmd_prop_plot <- ggplot(aasmd_prop, aes(Type, Value, fill = Type)) + geom_bar(stat = "identity") + ylab("Average Absolute Standardized Mean Difference") + xlab("Before vs After Matching") 


aasmd_prop_plot
```


## Step 3: Unsupervised Learning

### Step 3.1: Partitions Around Mediods(PAM) Clustering


First we need to remove columns before calculating dissimilarity: propensity, X, seqn, qsmk.

```{r}
library(dplyr)
diss_data <- smoking_data %>%
  select(sex, race, age, education.code, smokeintensity, smokeyrs, exercise, active, wt71)
```

```{r}
library(cluster)
#Generate gower distance 
gower_dist <- daisy(diss_data, metric = 'gower', type = list(factor = c(1, 2)))

#Convert gower distance into matrix
gower_mat <- as.matrix(gower_dist)

#Run PAM Clustering
pam_fit <- pam(gower_dist, diss = TRUE, k = 10)

#Add cluster assignment column to the diss_data set.
diss_data$cluster<- pam_fit$clustering

#Add the treamtent/control column back to the data:
diss_data$assignment <- smoking_data$qsmk
```

### Step 3.2: Comparing treamtent and control individuals in each cluster.

```{r}
library(Matching)


cluss_1 <- diss_data %>%
  filter(cluster == 1)
  

MatchBalance(assignment ~ sex +as.factor(race) + age + as.factor(education.code) + smokeintensity + smokeyrs + as.factor(exercise)+ active + wt71, data = cluss_1, nboots = 500)
  

summary(cluss_1)
```
